Forecasting 2024 US election using linear regression

### 2024-10-28 paper/paper.qmd
update with merged version

The data cleaning process involved filtering data to just high-quality
pollsters (i.e numeric_grade \>= 3.0) and filter to just Harris (i.e.
candidate_name == "Kamala Harris")

@@

> "Title of the Work" by Author Name, licensed under CC BY 4.0.

@@

After determining the number of respondents to be sampled from each
region, stratified sampling will be employed across key demographic
categories, including age, gender, race/ethnicity, and education level.
This approach ensures that the final sample accurately reflects the
diversity of the U.S. voting population by proportionally representing
each subgroup within every region. To mitigate potential non-response
bias, post-stratification weighting[^2] will be applied, correcting for
any imbalances caused by variations in response rates among different
demographic groups. The sample will be stratified based on several
critical demographic and geographic variables to guarantee proportional
representation of the U.S. voting population. Strata information will be
sourced from U.S. census data obtained through IPUMS USA
[@ruggles2024ipums].

[^2]: Post-stratification weighting adjusts survey data by applying
    weights to under- or over-represented demographic groups in the
    sample, ensuring the final results align with the true population
    distribution and reducing biases such as non-response bias.

@@

| **Region**    | **States**                                                                                                                                                 |
|-------------------------|-----------------------------------------------|
| **MIDWEST**   | Illinois, Indiana, Iowa, Kansas, Michigan, Minnesota, Missouri, Nebraska, North Dakota, Ohio, South Dakota, Wisconsin                                      |
| **Northeast** | Connecticut, Delaware, District of Columbia, Maine, Maryland, Massachusetts, New Hampshire, New Jersey, New York, Pennsylvania, Rhode Island, Vermont      |
| **SOUTH**     | Alabama, Arkansas, Florida, Georgia, Kentucky, Louisiana, Mississippi, North Carolina, Oklahoma, South Carolina, Tennessee, Texas, Virginia, West Virginia |
| **WEST**      | Alaska, Arizona, California, Colorado, Hawaii, Idaho, Montana, Nevada, New Mexico, Oregon, Utah, Washington, Wyoming                                       |

@@

| **Region**  | **Voter Turnout (%)** | **Total Ballots Cast** | **VEP**    | **Composite Measure Sampling Proportion** | **Sample Size** |
|------------|------------|------------|------------|------------|------------|
| **MIDWEST** | 69%                   | 35,134,960             | 50,932,439 | 0.214805579                               | 86              |
| **NW**      | 68%                   | 32,262,303             | 47,473,317 | 0.200216867                               | 80              |
| **SOUTH**   | 65%                   | 54,746,770             | 84,563,831 | 0.356644666                               | 143             |
| **WEST**    | 69%                   | 37,594,304             | 54,139,892 | 0.228332888                               | 91              |

: Regional Voting Data and Sample Size Allocation Based on 2020 Election
Turnout {#tbl-samplesize}

@@

    -   Not important
    -   1
    -   2
    -   3
    -   4
    -   5
    -   Very important

@@

1.  **Do you have any concerns or feedback regarding the survey,
    surveyor ,or entity?**\
    Your feedback is important to us and will help ensure transparency
    and trust in the research process.

>>>>

In this data-cleaning process, we focus on refining raw polling data to
enhance its quality and relevance for subsequent analysis. The dataset
is first loaded and cleaned using the janitor package to standardize
column names, ensuring consistency. We then filter the data to include
only relevant columns and remove any rows with missing values in
critical fields such as numeric_grade, pct, sample_size, and end_date.
Specifically, we isolate the polling data for Kamala Harris and Donald
Trump, applying a condition to keep only high-quality polls with a
numeric grade of 2 or higher since we find out that the mean of the
numeric grade is 2.175 while the median is 1.9.

For state-level polls, we address any placeholder states marked as "--"
by converting them to NA. A national poll indicator is created,
assigning a value of 1 for national polls and 0 for state-specific ones.
Dates are converted to a standard format using the lubridate package to
ensure correct handling in analyses. We calculate recency weights based
on how many days have passed since the poll ended, applying an
exponential decay to give more weight to recent polls. A sample size
weight is also implemented, capping the weights at a maximum of 2,300
responses to reflect the reliability of larger sample
sizes[@nyt_polling_averages]. Finally, the cleaned datasets for both
candidates are saved in CSV format for further modeling and analysis.
This systematic approach ensures that the data is accurate, complete,
and ready for insightful analysis.

## Analysis Data Summary Statistics

@tbl-analysisdata presents the summary statistics for each recorded
incident type for the years 2017 to 2022.

@@

@@

After determining the number of respondents to be sampled from each
region, stratified sampling will be employed across key demographic
categories, including age, gender, race/ethnicity, and education level.
This approach ensures that the final sample accurately reflects the
diversity of the U.S. voting population by proportionally representing
each subgroup within every region. To mitigate potential non-response
bias, post-stratification weighting\[\^2\] will be applied, correcting
for any imbalances caused by variations in response rates among
different demographic groups. The sample will be stratified based on
several critical demographic and geographic variables to guarantee
proportional representation of the U.S. voting population. Strata
information will be sourced from U.S. census data obtained through IPUMS
USA [@ruggles2024ipums].

\[\^2\]: Post-stratification weighting adjusts survey data by applying
weights to under- or over-represented demographic groups in the sample,
ensuring the final results align with the true population distribution
and reducing biases such as non-response bias.

@@

| **Region**    | **States**                                                                                                                                                 |
|--------------------------|----------------------------------------------|
| **MIDWEST**   | Illinois, Indiana, Iowa, Kansas, Michigan, Minnesota, Missouri, Nebraska, North Dakota, Ohio, South Dakota, Wisconsin                                      |
| **NORTHEAST** | Connecticut, Delaware, District of Columbia, Maine, Maryland, Massachusetts, New Hampshire, New Jersey, New York, Pennsylvania, Rhode Island, Vermont      |
| **SOUTH**     | Alabama, Arkansas, Florida, Georgia, Kentucky, Louisiana, Mississippi, North Carolina, Oklahoma, South Carolina, Tennessee, Texas, Virginia, West Virginia |
| **WEST**      | Alaska, Arizona, California, Colorado, Hawaii, Idaho, Montana, Nevada, New Mexico, Oregon, Utah, Washington, Wyoming                                       |

@@

| **Region**    | **Voter Turnout (%)** | **Total Ballots Cast** | **VEP**    | **Composite Measure Sampling Proportion** | **Sample Size** |
|------------|------------|------------|------------|------------|------------|
| **MIDWEST**   | 69%                   | 35,134,960             | 50,932,439 | 0.214805579                               | 86              |
| **NORTHEAST** | 68%                   | 32,262,303             | 47,473,317 | 0.200216867                               | 80              |
| **SOUTH**     | 65%                   | 54,746,770             | 84,563,831 | 0.356644666                               | 143             |
| **WEST**      | 69%                   | 37,594,304             | 54,139,892 | 0.228332888                               | 91              |

: Regional Voting Data and Sample Size Allocation calculated using
Composite Measure Sampling Proportion based on 2020 US Election regional
voter turnout (%) {#tbl-samplesize}

@@

    Not important
    -   1
    -   2
    -   3
    -   4
    -   5
    Very important

@@

1.  **Do you have any concerns or feedback regarding the survey,
    surveyor, or entity?**\
    Your feedback is important to us and will help ensure transparency
    and trust in the research process.
